---
title: "EDM - Deployment Exercise"
author: "Pablo Bolta, Jacinto Dobón y Jorge López"
date: "5/4/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(plotly)
library(reshape2)
library(lubridate)
library(randomForestSRC)
```

## 1.- One dimensional Partial Dependence Plot.
The partial dependence plot shows the marginal effect of a feature on the predicted outcome of a previously fit model. 
 
### EXERCISE:
Apply PDP to the regression example of predicting bike rentals. Fit a random forest approximation for the prediction of bike rentals (cnt). Use the partial dependence plot to visualize the relationships the model learned. Use the slides shown in class as model.  

### QUESTION:
Analyse the influence of days since 2011, temperature, humidity and wind speed on the predicted bike counts.

```{r exercise_1}
days <- read.csv("data/day.csv")

days$dteday <- as_date(days$dteday)
days_since <- select(days, workingday, holiday, temp, hum, windspeed, cnt)
days_since$days_since_2011 <- int_length(interval(ymd("2011-01-01"), days$dteday)) / (3600*24)
days_since$SUMMER <- ifelse(days$season == 3, 1, 0)
days_since$FALL <- ifelse(days$season == 4, 1, 0)
days_since$WINTER <- ifelse(days$season == 1, 1, 0)
days_since$MISTY <- ifelse(days$weathersit == 2, 1, 0)
days_since$RAIN <- ifelse(days$weathersit == 3 | days$weathersit == 4, 1, 0)
days_since$temp <- days_since$temp * 47 - 8
days_since$hum <- days_since$hum * 100
days_since$windspeed <- days_since$windspeed * 67
```


```{r rf_model_and_results}
model.rf <- rfsrc(cnt~., data=days_since)

results <- select(days_since, days_since_2011, temp, hum, windspeed, cnt)
len <- nrow(days_since)
for(c in names(results)[1:4])
{
  for(i in 1:len){
    r <- days_since
    r[[c]] <- days_since[[c]][i]
    pred <- predict(model.rf, r)$predicted
    results[[c]][i] <- sum(pred) / len
  }
}
```


```{r plot_results}
plot_1 <- ggplot(days_since, aes(x=days_since_2011, y=results$days_since_2011)) +
          geom_line() + 
          ylim(c(0,6000)) + 
          geom_rug(alpha=0.1, sides="b") + 
          ylab("Prediction") + 
          xlab("Days since 2011")

plot_2 <- ggplot(days_since, aes(x=temp, y=results$temp)) + 
          geom_line() + 
          ylim(c(0,6000)) + 
          geom_rug(alpha=0.1, sides="b") + 
          xlab("Temperature")

plot_3 <- ggplot(days_since, aes(x=hum, y=results$hum)) + 
          geom_line() + 
          ylim(c(0,6000)) + 
          geom_rug(alpha=0.1, sides="b") + 
          xlab("Humidity")

plot_4 <- ggplot(days_since, aes(x=windspeed, y=results$windspeed)) + 
          geom_line() + 
          ylim(c(0,6000)) + 
          geom_rug(alpha=0.1, sides="b") + 
          xlab("Wind speed")

subplot(plot_1, plot_2, plot_3, plot_4, 
          shareX = FALSE, 
          shareY = TRUE, 
          titleX = TRUE)
```

## 2.- Bidimensional Partial Dependency Plot.

### EXERCISE:
Generate a 2D Partial Dependency Plot with humidity and temperature to predict the number of bikes rented depending on those parameters.

> **BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot.**

Show the density distribution of both input features with the 2D plot as shown in the class slides. 

TIP: Use geom_tile() to generate the 2D plot. Set width and height to avoid holes. 

### QUESTION:
Interpret the results.

```{r}

```

## 3.- PDP to explain the price of a house.

### EXERCISE:
Apply the previous concepts to predict the price of a house from the database kc_house_data.csv. In this case, use again a random forest approximation for the prediction based on the features bedrooms, bathrooms, sqft_living, sqft_lot, floors and yr_built.Use the partial dependence plot to visualize the relationships the model learned.

> **BE CAREFUL: due to the size, extract a set of random samples from the BBDD before generating the data for the Partial Dependency Plot.** 

### QUESTION:
Analyse the influence of bedrooms, bathrooms, sqft_living and floors on the predicted price.

```{r}

```




